---
title: "Assignment2 - SQL Movies Data"
author: "John Ferrara"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task: Pull In a Custom SQL Data Table into An R Dataframe

This is the second file of the two files that is part if this assignment. The first file is a SQL file that creates the custom schema and table for use in R Studio.

## Connection to the SQL Server 

Connecting the the localhost SQL server in order to pull in my movies data. 

#### Step 1 - Importing and installing the packages needed
```{r sql_connect1}

#install.packages("RMySQL")
library(RMySQL)
```

#### Step 2 - Establish connection to MySQL database
```{r sql_connect2}
con <- dbConnect(RMySQL::MySQL(),
                 dbname = "assignment2_movies",
                 host = "127.0.0.1",
                 port = 3306,
                 user = "root",
                 password = "CUNYSPS")
```

#### Step 3 - Listing tables in the selected schema to confirm the connection
```{r sql_connect3}
dbListTables(con)
```

## Pulling in Data 
Now that the connection is established, pulling in the data to a dataframe. 

```{r dataframe}
movies_df <- dbReadTable(con, "movies")

head(movies_df)
```
