---
title: "FinalProject - DATA 606 & DATA 607"
author: "John Ferrara"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
#Loading libs
library(dplyr)
library(tidyverse)
library(glue)
library(lubridate)
library(jsonlite)
library(sf) # Dealing with Shapefiles
library(ggplot2)
```

## Abstract

## Introduction and Overview
As most know, in recent years the cost of housing has been outpacing wage growth^1^. Rising housing costs have led to a series of policy issues. While some impacts of high housing costs are obvious, others are less apparent. This analysis explores whether one such issue — the recovery of used syringes in public parks — can be linked to the housing crisis. In short, the question: Does housing impact this seemingly separate public health and safety issue of public, or improper, syringe disposal?

The null hypothesis for this study is that the number of syringes recovered from public parks by NYC Parks Department and other community-based non-profit staff, is an independent public health and safety issue. In other words, this issue is not influenced by a lack of housing units. The alternative hypothesis explored in this analysis is that the number of available housing units would have influence the number of used syringes recovered from public parks; that this dimension of public health risk, is reduced, at least in part, by housing policy.

It should be noted that these are large questions. Questions that are much more nuanced and complex than any single analysis can outline, so while this study seeks to demonstrate a link between two varying dimensions of public policy, it is by no means an attempt at a solution. Rather, the study seeks to determine if there is an overlap between the two issues using several sources of public data.

In order to properly articulate the case for such a connection, there are two main questions within this analysis. The first of the two is: Does the construction of housing units influence the number of homeless individuals? The second question is, does the size of an unhoused adult population influence the number of recovered syringes recovered in public parks?

## Data Sources
There are multiple data sources used in this study. Most, if not all, were sourced from NYC Open Data () a website that provides a multitude of public data sets generated by New York City government agencies. The data used in this analysis were:

**- Dataset 1: [NYC Parks Syringe Collection Data](https://data.cityofnewyork.us/Public-Safety/Summary-of-Syringe-Data-in-NYC-Parks/t8xi-d5wb/about_data)**

*Overview & Data Assumptions*

This dataset is from the NYC Parks Department. NYC Parks department staff, along with the staff of various community non-profit organizations, collect used syringes discarded improperly in public parks and log the totals. These collection totals are grouped by each Parks District, which is an internal administrative geographic boundary for the New York City Parks Department. Each row of data is the equivalent of a day’s total collection of syringes. Within this data set there are three total columns that outline the number of syringes collected. They are as follows:

* <u>Total Kiosk Syringes Collected:</u> This is the total number of syringes that are collected from the city’s safe disposal kiosks.

* <u>Total Ground Syringes Collected:</u>  This is the total number of syringes that are collected from the ground, or just generally found to be improperly disposed of.

* <u>Total Syringes Collected:</u> This is the total number of syringes from both the “ground” category and the “kiosk” category.
  
This analysis makes use of the total ground syringes collected, not the kiosk or ground syringes. While assumptions could be made about housed and unhoused population’s syringe disposal habits, the total number of syringes publicly disposed by either method should capture what is needed for this analysis.

*Source Format & Ingestion Method*

Sourced from CSV formatted files. The NYC Open Data API was used for iteratively ingesting this data from CSV structured formatting.


**- Dataset 2: [NYC Dept. City Planning Housing Database](https://data.cityofnewyork.us/Housing-Development/Housing-Database/6umk-irkx/about_data)**

*Overview & Data Assumptions*

This data is from the NYC Department of City Planning (DCP), the data contains information from the NYC Department of Buildings (DOB) for how many construction and demolition jobs are within a specific municipality-based geographies. For this analysis used at Community Districts for the main geographic boundary. The analysis leverages yearly counts of completed construction projects. The annual count data does not provide nuance between the three main types of construction projects included in the data (new buildings, major alterations, and demolitions), however, for the sake of this analysis, a larger number of projects is assumed to be targeted at enhancing living conditions and making new units available to the housing market. Annual counts from 2010 through 2020 by Community District were used in varying relative 5-year windows. How this was done is explained in detail within the methodology section.

The original download is a zipfile with multiple geographies, for this analysis the Community District boundary was used.

*Source Format & Ingestion Method*

The data source was downloaded in a zipfile containing multiple csv files for various geographic boundaries, the data was extracted, and a singular csv file containing housing units by Community District was ingested for processing.

**- Dataset 3: [NYC Parks Disticts](https://data.cityofnewyork.us/City-Government/NYC-Parks-Districts/mebz-ditc/about_data)**

*Overview & Data Assumptions*

This data set contained both the NYC Parks Department districts and Community Districts within one file. This geography data was processed so as to flatten the number of Community Districts that were listed for each respective Parks District. For example, if one row of data for Parks District X listed overlaps with Community Districts A, B and C. These three districts, in the raw data, are listed in one cell. This data was parsed so that the one row was extracted into three different rows, having one row for each Community District associated with the Parks District. THe data was essentially flattened long ways, the data was subsequently grouped to get a unique value count for Community Districts per Parks District. The processed table was used as a crosswalk to process the syringe numbers to obtain an estimate of syringe counts at the Community District level. As mentioned, the NYC Syringe Collection data only contained Parks districts, this crosswalk was joined into the syringe data in order to get syringe count estimates for Community Districts. This is discussed further in the methodology section.

*Source Format & Ingestion Method*

Read in directly from a single CSV URL.

**- Dataset 4: [NYC Dept. Homelessness Services Individual Census by Borough, Community District, and Facility Type](https://data.cityofnewyork.us/Social-Services/Individual-Census-by-Borough-Community-District-an/veav-vj3r/about_data)**

*Overview & Data Assumptions*

This dataset contains counts of individuals within the various types of shelters across New York City by community district for various reporting dates. The shelter types covered by this dataset include:

* <u>Adult Family (Commercial Hotel):</u> The population of adults in makeshift homeless shelters held within commercial hotels designated for adult families (i.e.m married couples with no children, a family with no children under the age of 21, or an unmarried couple who meets the DHS definition of a family unit).

* <u>Adult Family Shelter:</u> The population of adults in homeless shelters designated for adult families (i.e.m married couples with no children, a family with no children under the age of 21, or an unmarried couple who meets the DHS definition of a family unit).

* <u>Adult Shelter:</u> The population of adults in homeless shelters designated for single unhoused adults.

* <u>Adult Shelter (Commercial Hotel):</u> The population of adults in makeshift homeless shelters held within commercial hotels designated for single unhoused adults.

* <u>Family with Children (Commercial Hotel):</u> The population of individuals in makeshift homeless shelters held within commercial hotels designated for families with children.

* <u>Family with Children Shelter:</u> The population of individuals in homeless shelters designated for families with children.

For the sake of this analysis, only those shelters categories with adult-only numbers were counted, the assumption being is that intravenous drug users would be adults and not families with children. This means that the totals for Adult Family (Commercial Hotel), Adult Family Shelter, Adult Shelter, and Adult Shelter (Commercial Hotel) were used in the analysis for homeless populations within a community district. Those columns that had counts for family-only specific counts were not used.

*Source Format & Ingestion Method*

Sourced from JSON formatted files. The NYC Open Data API was used for iteratively ingesting this data from JSON structured formatting.

## Methodology
After ingesting the various data sets through their respective means, the data was processed. Each data set needed a unique series of processing steps in order to yield the finalized version of the data ultimately included in this analysis.

#### Syringe Data Processing
The Syringe data had totals for each park district on each date that the syringes were collected. For the sake of this analysis, the numbers were aggregated up to an annual collection total for each park district. Using the processed Community District to Park District crosswalk data, the final version of the crosswalk contained Community District, Parks District, and a count of distinct Community Districts for each parks district. The processed versions of both dfs were joined, and in order to yield an estimate for ground syringes recovered the Park District Syringe totals, which were native to the data, were divided by the number of community districts overlapping with each Parks District to generate a syringe estimate for each community district. 

#### Controlling for Varying Geographies 
As mentioned, the incongruities of the geographic units between the Community District-based housing and homeless datasets and the NYC Parks District syringe collection data set mandated the raw syringe totals be processed to yield a syringe estimate for Community District. Community Districts are a city-wide administrative boundary associated with the city’s community boards, while the parks districts are an internal NYC Parks Department administrative boundary. These two geographic boundaries do not have a 1:1 relationship with each other. There were instances of a singular community district overlapping with multiple parks districts, and vice versa. In order to approximate the number of syringes per community district, I simply divided the total number of syringes for each parks drastic by the total number of distinct Community Districts in each respective parks district. While this methodology is imperfect, the syringe collection data did not have Latitude and Longitude for each collected syringe or for specific collection sites, so a spatial join using shapefile geographies could not be completed.


#### Housing Data 
The housing data that was used was imported after downloading a zip file from 2023 Q4, and using one Community District-specific CSV from the batch of files. The CSV used contains annual counts fr the number of completed housing constructions projects fr each year from 2010 through 2023 by community district. For this analysis the assumption was made that the number of syringes collected in an area, provided that syringe collection is impacted by new housing units, would be a lagging indicator. With this mindset, the number of completed housing construction projects in a community district for five years prior to the year the syringes were collected, was the aggregate version of housing numbers used for the analysis In other words, for each year in the syringe collection data, the preceding five years of construction jobs were summed up and the average Year-Over-Year change in construction jobs for those years were calculated. For instance, for the total number of estimated syringes collected for a Community District in 2017, the metrics for housing-related construction jobs within that same community district would be the aggregate sum of all housing jobs from 2011 through 2016, as well as the average Year-Over-Year change in housing jobs for those years. This aggregation into two main columns for each annually summed community district row allowed for the housing numbers to be joined with the Syringe Collection data.

#### Homeless Population Data 
The Homeless shelter population data was pulled into the analysis in JSON format via NYC Open Data’s API. To process this data the sum of the following columns were used as an estimate for the total number of unhoused adults: adult_shelter, adult_shelter_comm_hotel, adult_family_shelter,and adult_family_comm_hotel. These columns were counts from various types of shelters throughout the city. The row-specific sums for these columns were then averaged for each community district on an annual basis, which was to control for multiple collections within a year. Once completed, the processed data was joined into the df that contained the syringes data and the housing construction data.


## Ingestion 

#### NYC Parks Syringe Collection Data
``` {r ingestion1}
## API Intake; Documentation states there are 35,155 rows of data. Iterate through these numbers for API intake ( Limited to 1,000 row chunks)
total_srg_rows <- 35155 #Manually Checked on Site
srg_endpnt <- "https://data.cityofnewyork.us/resource/t8xi-d5wb.csv"
srg_suffix <- '?$offset='

##Initial Pull
initial_pull <- read.table(srg_endpnt, header = TRUE, sep = ",", dec = ".")
offset <-nrow(initial_pull) # Should be the length of the API chunk limit. (e.g., 1000 rows)

##Iteratting through all of the chunks & Appending to empty DF
running_pulls <- data.frame()
for (i in seq(0, total_srg_rows, by = offset)) {
  if (i == 0){
    next
  }
  interim_endpoint <- paste(srg_endpnt,srg_suffix,i,sep = "")
  # print(interim_endpoint)
  interim_pull <- read.table(interim_endpoint, header = TRUE, sep = ",", dec = ".")
  running_pulls <- rbind(running_pulls,interim_pull)
}

##Putting all Srynge Pulls together 
raw_syrg <- rbind(initial_pull,running_pulls)

##confirming number of total rows
# print(nrow(initial_pull))
# print(nrow(running_pulls))
# print(nrow(raw_syrg))
```
#### NYC Dept. City Planning Housing Database
```{r ingestion2}
## Dataset 2: NYC DCP Housing Data (Broken Down by Multiple Geographic Area's; Choosing Community District Breakdown)

### Downloading the 23Q4 CSV Zip File from the [DCP Wesbite](https://www.nyc.gov/site/planning/data-maps/open-data/dwn-housing-database.page#housingdevelopmentproject).
### Selected "HousingDB_by_CommunityDistrict.csv" from zipped contents.
download.file("https://s-media.nyc.gov/agencies/dcp/assets/files/zip/data-tools/bytes/nychousingdb_23q4_csv.zip",
              destfile = "housing_zip.zip", mode = "wb")
unzip("housing_zip.zip", files = "HousingDB_by_CommunityDistrict.csv", exdir = tempdir())
housing_path <- file.path(tempdir(), "HousingDB_by_CommunityDistrict.csv")
raw_housing <- read.table(housing_path, header = TRUE, sep = ",", dec = ".")
```

#### NYC Parks Disticts
```{r ingestion3}
## Dataset 3: NYC Parks DIstricts to Community District Crosswalk (https://data.cityofnewyork.us/City-Government/NYC-Parks-Districts/mebz-ditc/about_data)
crsswlk<- read.table("https://data.cityofnewyork.us/resource/mebz-ditc.csv",header = TRUE, sep = ",", dec = ".")
```

#### NYC Dept. Homelessness Services Individual Census by Borough, Community District, and Facility Type
```{r ingestion4,warning=FALSE, message=FALSE}
## Pulling in Via JSON API 
# Endpoint details
total_rows <- 4325 #Manually found row count on site
hmls_endpnt <- "https://data.cityofnewyork.us/resource/veav-vj3r.json"
hmls_suffix <- "?$offset="

# Initial pull to determine chunk size
initial_pull <- fromJSON(paste0(hmls_endpnt, hmls_suffix, 0)) #Fetch the first chunk

offset <- nrow(initial_pull) # Number of rows in each chunk returned by the API
if (is.null(offset) || offset == 0) {
  stop("Initial pull returned no rows. Check the API endpoint or offset logic.")
}

# List to store the chunks
running_pulls <- list()
# Loop through offsets to fetch all chunks
for (i in seq(0, total_rows, by = offset)) {
  # Construct endpoint with offset
  interim_endpoint <- paste0(hmls_endpnt, hmls_suffix, i)
  # print(paste("Fetching data from:", interim_endpoint))
  
  # Fetch data from the current offset
  interim_pull <- fromJSON(interim_endpoint)
  ## Ensuring Zeros instead of nulls
  interim_pull[is.na(interim_pull)] <- 0
  ## Keeping the columns we want (Only Adult Counts)
  interim_pull<-interim_pull[c("report_date","borough","community_districts","adult_shelter",
                 "adult_shelter_comm_hotel", "adult_family_shelter","adult_family_comm_hotel")]
  # Check if the pull is empty (no more rows)
  if (length(interim_pull) == 0) {
    # print("No more data to fetch.")
    break
  }
  
  # Append the data to the list
  running_pulls[[length(running_pulls) + 1]] <- interim_pull
}

# Combine all fetched data into a single data frame
raw_hmls <- do.call(rbind, running_pulls)

# Confirming the total number of rows
# print(paste("Total rows fetched:", nrow(raw_hmls))) ## Shuld be4325

## Filling in Borough values based on Park District
raw_hmls_processed<- raw_hmls %>%
  filter(borough!="Westchester")%>%
  mutate(cd_boro_num = case_when(borough== "Bronx"~ "2",
              borough== "Queens" ~"4",
              borough== "Manhattan" ~"1",
              borough== "Brooklyn"~"3",
              borough=="Staten Island"~"5"),
         padded_cd = str_pad(community_districts, width = 2, side = "left", pad = "0"),
         data_year = year(ymd_hms(report_date))) %>%
  mutate(full_community_district = paste0(cd_boro_num,padded_cd))


## Data Type Conversion before Sum
raw_hmls_processed[, c("adult_shelter", "adult_shelter_comm_hotel", "adult_family_shelter","adult_family_comm_hotel")] <- lapply(raw_hmls_processed[, c("adult_shelter", "adult_shelter_comm_hotel", "adult_family_shelter","adult_family_comm_hotel")], function(x) {
  as.numeric(x)})

raw_hmls_processed$AdultHomelessCount <- rowSums(raw_hmls_processed[, c("adult_shelter", "adult_shelter_comm_hotel",
                                                        "adult_family_shelter","adult_family_comm_hotel")])
# limiting to the columns needed
raw_hmls_limited <- raw_hmls_processed[,c("data_year","borough","full_community_district","AdultHomelessCount")]
## Grouping for a 1:1 value on year and CD, averaging th values where there are multple.
hmls_final <- raw_hmls_limited %>%
  group_by(data_year,borough, full_community_district) %>%
  summarise(
    avg_homeless_count = mean(AdultHomelessCount, na.rm=TRUE) )

hmls_final<-hmls_final %>% rename("communitydistrict"="full_community_district",
                                  "year"="data_year")
```

## Processing 

#### Starting with Syringe Data and Geographic Crosswalk
*Flattening the crosswalk longways. Multiple Community District Values in for a singular Parks District Row. Once flattened, the table can be used to  for Park Districts to Community Districts to enrich the Syringe Data with Community District.*
``` {r processing1}
## Firstly, limiting to the columns i need.
crsswlk_lim <- crsswlk[,c("borough","communityboard","parkdistrict")] 

## The NYC Parks DIstricts are not the same (1:1) as the Community Districts. Multiple community Districts for each park district. Need to flatten.
crsswlk_parsed <- data.frame()
for (i in 1:nrow(crsswlk_lim)) {
  row <- crsswlk_lim[i, ]
  cd_raw_char <- as.character(row$communityboard)
  split_values <- substring(cd_raw_char, seq(1, nchar(cd_raw_char), 3), seq(3, nchar(cd_raw_char), 3))

  expanded_df <- data.frame(
    communitydistrict = split_values,
    parkdistrict = row$parkdistrict,
    brough = row$borough
  )
  crsswlk_parsed <- rbind(crsswlk_parsed, expanded_df)}

# Sorting DF
crsswlk_parsed <- crsswlk_parsed %>% arrange(communitydistrict)
```

*Grouping Data to get Total Counts of Community District per Parks District*
```{r processing2}
## Getting Denominator for Aggregate Park District Syringe Totals; The number of Community Districts within each Park District. 
crsswlk_div_num_cd <- crsswlk_parsed %>% 
                    group_by(parkdistrict) %>%
                    summarise(cd_count_for_pd = n_distinct(communitydistrict))

## Crosswalk Final; CD Counts added, and data grouped. 
crosswalk_final <- merge(crsswlk_parsed,crsswlk_div_num_cd, by = "parkdistrict", all = FALSE)
```

*Further Cleaning, Null Removal, and Limiting to what is Needed*
``` {r processing3}
## Removing those entries that have no value for a park district 
# print(nrow(raw_syrg)) #35,155
syringe_lim <- raw_syrg %>% filter(district != "") 

## Filling in Borough values based on Park District
syringe_lim<- syringe_lim %>%
 mutate(borough = ifelse(
    borough == "",  # Check if `category` is blank
    case_when(
      substr(district, 1, 1) == "X" ~ "Bronx",
      substr(district, 1, 1) == "M" ~ "Manhattan",
      substr(district, 1, 1) == "Q" ~ "Queens",
      substr(district, 1, 1) == "B" ~ "Brooklyn",
      substr(district, 1, 1) == "R" ~ "Staten Island",
    ),
    borough
  ))

# print(nrow(syringe_lim)) #35,060 (95 Rows Removed)

### Limiting the raw syringe data to the columns i need for my analysis
syringe_lim <- syringe_lim[,c("year","group","location","borough","district","property_type","ground_syringes","kiosk_syringes","total_syringes")]
```

*Syringe Approximation by Community District*
``` {r proccessing4, warning=FALSE, message=FALSE}
syringe_grouped <- syringe_lim %>% 
  group_by(year,borough, district) %>%
  summarise(
    total_syringes = sum(total_syringes, na.rm=TRUE),
    ground_syringes= sum(ground_syringes, na.rm=TRUE)
  )
# print(head(syringe_grouped))

#renaming columns as needed 
syringe_grouped <- syringe_grouped %>% rename(parkdistrict = district)

## Looking at min and max years
# print(min(syringe_grouped$year))#2017
# print(max(syringe_grouped$year))#2024

syringe_grouped_enr <- merge(syringe_grouped, crosswalk_final, by = "parkdistrict", all = FALSE)
syringe_massaged <- syringe_grouped_enr %>% mutate(ttl_syring_est_interim =  total_syringes / cd_count_for_pd)

syringe_final <- syringe_massaged %>% 
  group_by(year,borough,communitydistrict) %>%
  summarise(total_syringe_ests = sum(ttl_syring_est_interim, na.rm=TRUE) )
```

#### Starting with Housing Data
*Summarizing data for total new Units by CD from the DOB for various rolling 5 years periods for each of the years within the syringe data*
``` {r processing5}
## Limiting Columns for is needed for analysis
raw_housing<-raw_housing[,c("commntydst","comp2010","comp2011","comp2012","comp2013","comp2014","comp2015","comp2016","comp2017",
               "comp2018","comp2019","comp2020","comp2021","comp2022","comp2023")]

## Summarizing data for totals for various rolling 5 years periods for each of the years within the syringe data
###Syringe Rolling Years
housing_agg_df <- data.frame()
for (i in 1:nrow(syringe_grouped)) {
  row <- syringe_grouped[i, ]
  # row$year
  year_range <- as.character(as.integer(seq(from = (row$year-5), to = (row$year-1), length.out = 5)))
  # print(year_range)
  temp_housing <- raw_housing[,c(
                                 "commntydst",
                                 glue("comp{year_range[1]}"),
                                 glue("comp{year_range[2]}"),
                                 glue("comp{year_range[3]}"),
                                 glue("comp{year_range[4]}"),
                                 glue("comp{year_range[5]}"))]
  temp_housing$syringe_yr <-row$year
  temp_housing <- temp_housing %>% rename("comp_yr1" = glue("comp{year_range[1]}"),
                                          "comp_yr2" = glue("comp{year_range[2]}"),
                                          "comp_yr3" = glue("comp{year_range[3]}"),
                                          "comp_yr4" = glue("comp{year_range[4]}"),
                                          "comp_yr5" = glue("comp{year_range[5]}"))
  housing_agg_df <- rbind(housing_agg_df, temp_housing)
}

#Looking at result
#print(head(housing_agg_df))

# Relative Yearly Sums for individual years within 5 yr window
housing_sums <- housing_agg_df %>%
  mutate(housingsum = comp_yr1 + comp_yr2 + comp_yr3 + comp_yr4 + comp_yr5,
    yoy_yr2 = ifelse(comp_yr1 == 0, NA, (comp_yr2 - comp_yr1) / comp_yr1 * 100),
    yoy_yr3 = ifelse(comp_yr2 == 0, NA, (comp_yr3 - comp_yr2) / comp_yr2 * 100),
    yoy_yr4 = ifelse(comp_yr3 == 0, NA, (comp_yr4 - comp_yr3) / comp_yr3 * 100),
    yoy_yr5 = ifelse(comp_yr4 == 0, NA, (comp_yr5 - comp_yr4) / comp_yr4 * 100)) %>%
  mutate(
    avg_yoy_change = rowMeans(select(., yoy_yr2, yoy_yr3, yoy_yr4, yoy_yr5), na.rm = TRUE)
  )

housing_final <-housing_sums[,c("syringe_yr","commntydst","housingsum","avg_yoy_change")] %>%
                rename("communitydistrict"="commntydst")
#housing_final
```

#### Putting it all Together; the Final DF
``` {r processing6}
## Joining in the homeless data to the syringe data (full join for maps. Will limit for regressions later.)
syringe_homeless_fnl<-merge(syringe_final,hmls_final, by = c('year',"borough","communitydistrict"), all = TRUE)

housing_final_join <-housing_final %>% rename("year"="syringe_yr")

# checking data types 
syringe_homeless_fnl$year <- as.integer(syringe_homeless_fnl$year)
syringe_homeless_fnl$communitydistrict <- as.integer(syringe_homeless_fnl$communitydistrict)
housing_final_join$year <- as.integer(housing_final_join$year)
housing_final_join$communitydistrict <- as.integer(housing_final_join$communitydistrict)

housing_syringe_df <- merge(syringe_homeless_fnl, 
                           housing_final_join, 
                           by= c("year","communitydistrict"),all=TRUE)

housing_syringe_df<-housing_syringe_df%>%distinct()
# housing_syringe_df

data_final <- housing_syringe_df[,c("year","borough","communitydistrict",
                                    "total_syringe_ests","avg_homeless_count",
                                    "housingsum","avg_yoy_change")]%>%
  rename("prev5_avg_yoy_change"="avg_yoy_change")
# data_final

```

## Analysis

Firstly, let's take a look at the data itself and see what the processed data shows us about these three dimensions of data. 

### Basic Summary Statistics 
```{r sum_stat}

## Number of Rows in finalized Data
print(nrow(data_final)) #570

## Year Coverage for variables
#Syringes
unique(data_final$year[!is.na(data_final$total_syringe_ests)])
#Syringe Counts have Values from 2017 through 2024

#Homeless Shelter Pops
unique(data_final$year[!is.na(data_final$avg_homeless_count)])
# 2018 through 2024

#Construction Housing Projecrs ( this should match the Syringe Count b/c of processing)
unique(data_final$year[!is.na(data_final$housingsum)])
# 2017 through 2024


## Means for Variables (Need to remove NUlls)
print(mean(data_final$total_syringe_ests,  na.rm = TRUE)) #4,875.857

print(mean(data_final$avg_homeless_count,  na.rm = TRUE)) #364.3451

print(mean(data_final$housingsum,  na.rm = TRUE)) # 1,624.679

print(mean(data_final$prev5_avg_yoy_change,  na.rm = TRUE)) #232.8672

## Std Dev for Variables
print(sd(data_final$total_syringe_ests,  na.rm = TRUE)) #13,780.57

print(sd(data_final$avg_homeless_count,  na.rm = TRUE)) #363.9017

print(sd(data_final$housingsum,  na.rm = TRUE)) # 2,100.204

print(sd(data_final$prev5_avg_yoy_change,  na.rm = TRUE)) #2,072.639

```

### Prepping for Map Visuals
```{r  geojson_map_prep}
## Dealing with GeoJSON file in order to map data by Community District
cd_geojson <- tempfile(fileext = ".geojson")
download.file("https://data.cityofnewyork.us/api/geospatial/yfnk-k7r4?method=export&format=GeoJSON",
              cd_geojson)
## Formatting the Community DIstrict Shapefile as needed to join with the final df.
cd_sf <- read_sf(cd_geojson)
cd_sf <- cd_sf %>% rename("communitydistrict"="boro_cd")
cd_sf$communitydistrict <- as.integer(cd_sf$communitydistrict)

### Looping through all the years in the "data_final" df to do the join for geom, to keep the blank geoms needed for full map
adjusted_geom = data.frame()
for (y in unique(data_final$year)){
  lim_data <-data_final %>% filter(year==y)
  # print(length(unique(lim_data$communitydistrict)))
  # Left Joining for each year, so we have Geom for all CD, even if not in the data.
  temp_df <- cd_sf %>%
    left_join(lim_data, by = "communitydistrict")
  temp_df$year <- y
  adjusted_geom<-rbind(adjusted_geom,temp_df)
}

#Ensuring the borough Values arent null for those CD that are not in the Housing/Syringe Data
adjusted_geom <- adjusted_geom %>% mutate(borough = ifelse(is.na(borough),
                                    case_when(
                                      substr(as.character(communitydistrict), 1, 1) == "2" ~ "Bronx",
                                      substr(as.character(communitydistrict), 1, 1) == "1" ~ "Manhattan",
                                      substr(as.character(communitydistrict), 1, 1) == "4" ~ "Queens",
                                      substr(as.character(communitydistrict), 1, 1) == "3" ~ "Brooklyn",
                                      substr(as.character(communitydistrict), 1, 1) == "5" ~ "Staten Island",
                                      ),
                                    borough
                                    ))
```

#### Mapping To Visually See The Data in a Geospatial Context

**Syringe Data by Community District**

Looking at the year by year break down for each syringe collection year in the data, one can identify that there are a decent number of nulls within the data. This could be from either syringes not being found by Parks staff in the vast majority of the city, or it could have to do with when and how the data started being collected. In the earlier years the syringe counts are limited to the South Bronx and small sections of Manhattan, but as the years progress additional parts of the city register a having a syringes collected.

```{r mapping1, echo=FALSE,  warning=FALSE, message=FALSE}
#Defining the Palette 
palette <- scale_fill_viridis_c(option = "inferno", na.value = "grey85", limits = range(adjusted_geom$total_syringe_ests, na.rm = TRUE),)

## Mapping Syringe Collection for Each Year
ggplot(adjusted_geom) +
  geom_sf(aes(fill = total_syringe_ests), color = "white",  na.rm = FALSE) +
  theme_void() + # No Background
  palette+
  theme(plot.title = element_text(hjust = 0.5, size = 12),
        legend.position = c(0.89, 0.04),
        legend.justification = "bottom") +
  labs(fill = "Est. Syringes Count", title = "Annual Est. Syringes Recovered in NYC Parks by Comm. Dist.") +
  facet_wrap(~year, ncol = 3) 
#saving
# ggsave("SyringesMap.png", width = 8, height = 8, bg='white')

```

**Total Number of Housing Related Construction Projects for 5 Years Previous by Community District**

For the total number of housing related completed construction projects registered with DOB, in the early years the high development zones are limited to Northern Brooklyn and parts on the lower end of Manhattan. As the years progress, the number of projects through out the city, more specifcally in Brooklyn and Queens, increase.

```{r mapping2, echo=FALSE,  warning=FALSE, message=FALSE}
palette <- scale_fill_viridis_c(option = "inferno", na.value = "grey85", limits = range(adjusted_geom$housingsum, na.rm = TRUE),)

## Mapping Housing Construction (SUM) Numbers for Each Year
ggplot(adjusted_geom) +
  geom_sf(aes(fill = housingsum), color = "white",  na.rm = FALSE) +
  theme_void() + # No Background
  palette+
  theme(plot.title = element_text(hjust = 0.5, size = 12),
        legend.position = c(0.89, 0.04),
        legend.justification = "bottom") +
  labs(fill = "Completed Project Count", title = "Total Number of Housing Projects for Preceeding 5 Yrs by Comm. Dist.") +
  facet_wrap(~year, ncol = 3) 
#saving
# ggsave("SumHousingMap.png", width = 8, height = 8, bg='white')

```

**Average YOY Change for Housing Related Construction Projects 5 Years Previous by Community District**

The average YOY change for completed housing related construction project is faily consistant through out the entire city through out the years in the data. However, the areas of what looks like Sunset Park & Park Slope in Brooklyn exprienced noticably high YOY change in projecs from 2017 through 2020. 

```{r mapping3, echo=FALSE,  warning=FALSE, message=FALSE}
palette <- scale_fill_viridis_c(option = "inferno", na.value = "grey85", limits = range(adjusted_geom$prev5_avg_yoy_change, na.rm = TRUE),)

## Mapping Housing Construction (SUM) Numbers for Each Year
ggplot(adjusted_geom) +
  geom_sf(aes(fill = prev5_avg_yoy_change), color = "white",  na.rm = FALSE) +
  theme_void() + # No Background
  palette+
  theme(plot.title = element_text(hjust = 0.5, size = 12),
        legend.position = c(0.89, 0.04),
        legend.justification = "bottom") +
  labs(fill = "Avg. YOY Percent Change", title = "5 Yr Avg. Pct YOY Change Completed Housing Construction Projects by Year by Comm. Dist.") +
  facet_wrap(~year, ncol = 3) 
#saving
# ggsave("HousingYOYChangeMap.png", width = 8, height = 8, bg='white')

```

**Annual Average Adult Homeless Population Based On Shelter Counts**

Lastly, for the average annual adult homeless population, there is no data for 2017, but once data is collected for 2018 there are many different parts of the city that have fairly high average adult homeless populations. There doesnt seem to be any particular trend visible just by looking at the data. However, it should be noted, that while this data is useful for gauging the homeless population in the city, at a Community District geographic boundary level, the means by which the data is collected - taking counts at homeless shelters of different kinds - implicity causes the homeless population to be higher in areas that have hotels or shelters designated for that population. This may not necessarily be an issue, but it is something to keep in mind when comparing this variable to others. 

```{r mapping4, echo=FALSE,  warning=FALSE, message=FALSE}
palette <- scale_fill_viridis_c(option = "inferno", na.value = "grey85", limits = range(adjusted_geom$avg_homeless_count, na.rm = TRUE),)

## Mapping Housing Construction (SUM) Numbers for Each Year
ggplot(adjusted_geom) +
  geom_sf(aes(fill = avg_homeless_count), color = "white",  na.rm = FALSE) +
  theme_void() + # No Background
  palette+
  theme(plot.title = element_text(hjust = 0.5, size = 12),
        legend.position = c(0.89, 0.04),
        legend.justification = "bottom") +
  labs(fill = "Avg. Adult Homeless Population", title = "Annual Average of Adult Homeless Population by Comm. Dist.") +
  facet_wrap(~year, ncol = 3) 
#saving
# ggsave("HomelessPopMap.png", width = 8, height = 8, bg='white')

```

#### Removing Nulls from Finalized DF
``` {r null_cleaning}
#Removing the Nulls from the final data df in preparation for regression
data_final_nonull <- na.omit(data_final)
# data_final_nonull
```


### Linear Regressions w/ Community District Geographies
As outlined in the introduction, there are two parts to this analysis. The first is to see if there is a correlation between housing-related construction projects and the homeless population. The second is to see if the homeless population correlates with the amount of syringes found in NYC parks and other public safe disposal areas. This first attempt at regresssions will maintain a larger number of data points contrained by community district. However, the varying factors that influence various community districts may influence the regression results obfuscating any substantial correlation. Factors such as high demand, gentrifying areas of the city having more construction projects associated with their geographic boundary, while intravenous drug use is more likely to be concentrated in poorer, less "in-demand" community districts. Additionally, addtionally housing units through out the city may not properly be shown to reduce homelessnes due to the homeless population data being sourced from shelters. Shelters are not evenly distributed throughout the city, making the numbers at the community district level skewed for the overall numbers.

#### QUESTION 1: Does the Number of Housing Related Construction Projects for the Previous 5 Years influence the Size of Adult Homeless Population?

To answer this question, a linear regression analysis will look at the relationship between the annual average for adult homeless shelter populations and the number of completed housing construction jobs on record wit hthe DOB. For the housing numbers, we will look at the average annual year-over-year change in housing construction projects for the five years previous to the homeless population year, as well as the total number of cumulative housing construction projects completed in for the five years previous to the homeless population year. 

**Relationship Between Adult Homeless Population and Completed Housing Construction Projects for Preceeding 5 Years at the Community District Level**
``` {r reg1a, echo=FALSE,  warning=FALSE, message=FALSE}

## Ground Syringes to Housing SUm
m1a <- lm(avg_homeless_count ~ housingsum, data = data_final_nonull)
print(summary(m1a))
```

```{r reg1a_plot, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data = data_final_nonull, aes(x = housingsum, y = avg_homeless_count)) +
geom_point() +
stat_smooth(method = "lm", se = FALSE)+
  labs(x="Total Number of Completed Housing Related Construction Projects for Preceeding 5 Years",
       y="Annual Average of Adult Homeless Population",
       # title="Relationship Between Adult Homeless Population and Completed Housing Construction Projects for Preceeding 5 Years at the Community District Level"
       )
```


### Model Is Statisitcally Significant So Checking Model Validity
**Model Linearity Check**
``` {r m1a_check1} 
# Linearity Check
ggplot(m1a, aes(x=.fitted, y=.resid)) + 
  geom_point()+
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Fitted values",y ="Residuals")
## Not completely Zero, Seems to be correlation in lower left. However line seems to have Zero Slope
```

**Model Residual Distribution Check 1 **
``` {r m1a_check2}
# Residual Dist.
ggplot(data = m1a, aes(x = .resid)) +geom_histogram(binwidth = 10) + xlab("Residuals")
## Mostly Normal, Right Skewed
```

**Model Residual Distribution Check 2**
``` {r m1a_check3}
#Variablitiy of Constant
ggplot(data = m1a, aes(sample = .resid)) +stat_qq()
## Slight Deviation b/c of right skew, but passes for normal.
```


**Relationship Between Logged Adult Homeless Population and Logged Completed Housing Construction Projects for Preceeding 5 Years at the Community District Level**
``` {r log_reg1a, echo=FALSE, warning=FALSE, message=FALSE}

## Ground Syringes to Housing SUm
lm1a <- lm(log(avg_homeless_count+1) ~ log(housingsum+1), data = data_final_nonull)
print(summary(lm1a))
```

```{r logreg1a_plot, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data = data_final_nonull, aes(x = log(housingsum+1), y = log(avg_homeless_count+1))) +
geom_point() +
stat_smooth(method = "lm", se = FALSE)+
  labs(x="Logged Total Number of Completed Housing Related Construction Projects for Preceeding 5 Years",
       y="Logged Annual Average of Adult Homeless Population",
       # title="Relationship Between Logged Adult Homeless Population and Logged Completed Housing Construction Projects for Preceeding 5 Years at the Community District Level"
       )
```

**Relationship Between Adult Homeless Population and Average YOY Percent Change in Completed Housing Related Construction Projects for Preceeding 5 Years for Preceeding 5 Years at the Community District Level**

``` {r reg1b, echo=FALSE, warning=FALSE, message=FALSE}
## Ground Syringes to Housing SUm
m1b <- lm(avg_homeless_count ~ prev5_avg_yoy_change, data = data_final_nonull)
print(summary(m1b))
```

```{r reg1b_plot, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data = data_final_nonull, aes(x = prev5_avg_yoy_change, y = avg_homeless_count)) +
geom_point() +
stat_smooth(method = "lm", se = FALSE)+
  labs(x="Avg. YOY Pct Change in Completed Housing Related Constr. Projects for Preceeding 5 Years",
       y="Annual Average of Adult Homeless People",
       # title="Relationship Between Adult Homeless Population and Average YOY Percent Change in Completed Housing Related Construction Projects for Preceeding 5 Years for Preceeding 5 Years at the Community District Level"
       )
```

**Relationship Between Logged Adult Homeless Population and Logged Completed Housing Related Construction Projects for Preceeding 5 Years at the Community District Level**
``` {r log_reg1b, echo=FALSE, warning=FALSE, message=FALSE}

## Ground Syringes to Housing SUm
lm1b <- lm(log(avg_homeless_count+1) ~ log(prev5_avg_yoy_change+1), data = data_final_nonull)
print(summary(lm1b))
```

```{r logreg1b_plot, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data = data_final_nonull, aes(x = log(prev5_avg_yoy_change+1), y = log(avg_homeless_count+1))) +
geom_point() +
stat_smooth(method = "lm", se = FALSE)+
  labs(x="Logged Avg. YOY Pct Change in Completed Housing Related Constr. Projects for Preceeding 5 Years",
       y="Logged Annual Average of Adult Homeless People by CD",
       #title="Relationship Between Logged Adult Homeless Population and Logged Completed Housing Related Construction Projects for Preceeding 5 Years at the Community District Level"
       )
```

#### QUESTION 2: Does the size of the Adult Homeless Population influence the Number of Used Syringes recovered by NYC Parks? 

**Relationship Between Adult Homeless Population and Est. Total Number of Syringes Recovered at the Community District Level**
``` {r reg2, echo=FALSE, warning=FALSE, message=FALSE}

## Ground Syringes to Avg. Annual Housing Increase
m2 <- lm(total_syringe_ests ~ avg_homeless_count, data = data_final_nonull)
print(summary(m2))
```

```{r reg2_plot, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data = data_final_nonull, aes(x = avg_homeless_count, y = total_syringe_ests)) +
geom_point() +
stat_smooth(method = "lm", se = FALSE)+
  labs(x="Est. Total Number of Used Syringes Collected",
       y=" Annual Average Adult Homeless Population",
       # title="Relationship Between Adult Homeless Population and Est. Total Number of Syringes Recovered at the Community District Level"
       )
```

**Relationship Between Logged Adult Homeless Population and Logged Est. Total Number of Syringes Recovered at the Community District Level**
``` {r logreg2, echo=FALSE, warning=FALSE, message=FALSE}

## Ground Syringes to Avg. Annual Housing Increase
lm2 <- lm(log(total_syringe_ests+1) ~ log(avg_homeless_count+1), data = data_final_nonull)
print(summary(lm2))
```

```{r logreg2_plot, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data = data_final_nonull, aes(x = log(avg_homeless_count+1), y = log(total_syringe_ests+1))) +
geom_point() +
stat_smooth(method = "lm", se = FALSE)+
    labs(x="Logged Est. Total Number of Used Syringes Collected",
       y=" Logged Annual Average of Adult Homeless People",
       # title="Relationship Between Logged Adult Homeless Population and Logged Est. Total Number of Syringes Recovered at the Community District Level"
       )
```

### Linear Regressions with Aggregate Borough Geographies
After performing the regression analysis above, the results were unremarkable. There was one statisticly significant relationship - the direct correlation with the number of construction related housing projects and the size of the adult homeless population. Due to potential issues with the data at the Community District level, such as an unequal distribution of parks where syrings can be found, specific types of zoning limiting where shelters exist implicitly skewing homeless populations, or housing projects being limited to select areas. I decided to aggregate to the borough level instead. Generalizing the geographic limitations may show relationships more clearly than the more granular boundaries. The main draw back to this aggregation is a reduction in datapoints, however it is worth taking a look.

#### Aggregating the data to the borough level.

**Syringe Borough Aggregation**
``` {r borough_agg1, warning=FALSE, message=FALSE}
syringe_grouped_boro <- syringe_lim %>% 
  group_by(year,borough) %>%
  summarise(
    total_syringes = sum(total_syringes, na.rm=TRUE)
    )
```

**Homeless Borough Aggregation**
```{r boroughagg2, warning=FALSE, message=FALSE}

raw_hmls_limited_boro <- raw_hmls_processed[,c("report_date","data_year","borough","AdultHomelessCount")]

## Grouping by report date to sum up boro total for report date.
hmls_grpd_boro <- raw_hmls_limited_boro %>%
  group_by(report_date,data_year,borough) %>%
  summarise(
    boro_sum_homeless_count = sum(AdultHomelessCount, na.rm=TRUE))

## Averaging for each boto for each year
hmls_final_boro <- hmls_grpd_boro %>%
  group_by(data_year,borough) %>%
  summarise(
    avg_homeless_count = mean(boro_sum_homeless_count, na.rm=TRUE))

hmls_final_boro<-hmls_final_boro %>% rename("year"="data_year")

```

**Housing Borough Agg**
```{r boroughagg3, warning=FALSE, message=FALSE}
housing_agg_df_boro <- housing_agg_df %>%
  mutate(borough = case_when(
    substr(as.character(commntydst), 1, 1) == "2" ~ "Bronx",
    substr(as.character(commntydst), 1, 1) == "4" ~ "Queens",
    substr(as.character(commntydst), 1, 1) == "3" ~ "Booklyn",
    substr(as.character(commntydst), 1, 1) == "5" ~ "Staten Island",
    substr(as.character(commntydst), 1, 1) == "1" ~ "Manhattan"))

housing_agg_df_boro_grpd<- housing_agg_df_boro %>%
  group_by(syringe_yr, borough) %>%
  summarize(
    comp_yr1 = sum(comp_yr1, na.rm=TRUE),
    comp_yr2 = sum(comp_yr2, na.rm=TRUE),
    comp_yr3 = sum(comp_yr3, na.rm=TRUE),
    comp_yr4 = sum(comp_yr4, na.rm=TRUE),
    comp_yr5 = sum(comp_yr5, na.rm=TRUE)
  )

housing_sums_boro <-   housing_agg_df_boro_grpd %>%
  mutate(housingsum = comp_yr1 + comp_yr2 + comp_yr3 + comp_yr4 + comp_yr5,
    yoy_yr2 = ifelse(comp_yr1 == 0, NA, (comp_yr2 - comp_yr1) / comp_yr1 * 100),
    yoy_yr3 = ifelse(comp_yr2 == 0, NA, (comp_yr3 - comp_yr2) / comp_yr2 * 100),
    yoy_yr4 = ifelse(comp_yr3 == 0, NA, (comp_yr4 - comp_yr3) / comp_yr3 * 100),
    yoy_yr5 = ifelse(comp_yr4 == 0, NA, (comp_yr5 - comp_yr4) / comp_yr4 * 100))

housing_sums_boro$avg_yoy_change <- rowMeans(housing_sums_boro[, c('yoy_yr2', 'yoy_yr3', 'yoy_yr4', 'yoy_yr5')])

housing_final_boro <- housing_sums_boro[,c("syringe_yr","borough","housingsum","avg_yoy_change")] %>%
                rename("year"="syringe_yr")

```

**Joining Aggregated Data Sets Together for Borough Level Geography**
```{r boro_joins, warning=FALSE, message=FALSE}
# first join
 syringe_homeless_boro<- merge(syringe_grouped_boro,hmls_final_boro, by = c('year',"borough"), all = TRUE)

# checking data types 
syringe_homeless_boro$year <- as.integer(syringe_homeless_boro$year)
housing_final_boro$year <- as.integer(housing_final_boro$year)

#second join
housing_syringe_df_boro <- merge(syringe_homeless_boro, 
                           housing_final_boro, 
                           by= c("year","borough"),all=TRUE)

#To keep more data points with the Agg, making the NA syrings values zero
housing_syringe_df_boro <- housing_syringe_df_boro %>%
    mutate(total_syringes = ifelse(is.na(total_syringes), 0, total_syringes))

# Omitting the remaining nulls
data_final_boro_nonulls<- na.omit(housing_syringe_df_boro)
```

#### QUESTION 1: Does the Number of Housing Related Construction Projects for the Previous 5 Years influence the Size of Adult Homeless Population?

**Relationship Between Adult Homeless Population and Total Number of Completed Housing Related Construction Projects for Preceeding 5 Years at the Borough Level**
``` {r agg_reg1a, echo=FALSE, warning=FALSE, message=FALSE}

## Ground Syringes to Housing SUm
agg_m1a <- lm(avg_homeless_count ~ housingsum, data = data_final_boro_nonulls)
print(summary(agg_m1a))
```

```{r agg_reg1a_plot, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data = data_final_boro_nonulls, aes(x = housingsum, y = avg_homeless_count)) +
geom_point() +
stat_smooth(method = "lm", se = FALSE)+
        labs(x="Total Number of Completed Housing Related Constr. Projects for Preceeding 5 Years",
       y="Annual Average of Adult Homeless Population",
       # title="Relationship Between Adult Homeless Population and Total Number of Completed Housing Related Construction Projects for Preceeding 5 Years at the Borough Level"
       )
```

**Relationship Between Logged Adult Homeless Population and Logged Total Number of Completed Housing Related Construction Projects for Preceeding 5 Years at the Borough Level**
``` {r agg_log_reg1a, echo=FALSE, warning=FALSE, message=FALSE}
## Ground Syringes to Housing SUm
agg_lm1a <- lm(log(avg_homeless_count+1) ~ log(housingsum+1), data = data_final_boro_nonulls)
print(summary(lm1a))
```

```{r agg_logreg1a_plot, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data = data_final_boro_nonulls, aes(x = log(housingsum+1), y = log(avg_homeless_count+1))) +
geom_point() +
stat_smooth(method = "lm", se = FALSE)+
      labs(x="Logged Total Number of Completed Housing Rltd Constr. Projects for Preceeding 5 Years",
       y=" Logged Annual Average of Adult Homeless Population",
       # title="Relationship Between Logged Adult Homeless Population and Logged Total Number of Completed Housing Related Construction Projects for Preceeding 5 Years at the Borough Level"
       )
```

**Relationship Between Adult Homeless Population and Average YOY Percent Change in Completed Housing Related Construction Projects for Preceeding 5 Years at the Borough Level**

``` {r agg_reg1b, echo=FALSE, warning=FALSE, message=FALSE}

## Ground Syringes to Housing SUm
agg_m1b <- lm(avg_homeless_count ~ avg_yoy_change, data = data_final_boro_nonulls)
print(summary(agg_m1b))
```

```{r agg_reg1b_plot, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data = data_final_boro_nonulls, aes(x = avg_yoy_change, y = avg_homeless_count)) +
geom_point() +
stat_smooth(method = "lm", se = FALSE)+
  labs(x="Average YOY Pct. Chng in Completed Housing Rltd Constr. Projects for Preceeding 5 Years",
       y="Annual Average Adult Homeless Pppulation",
       # title="Relationship Between Adult Homeless Population and Average YOY Percent Change in Completed Housing Related Construction Projects for Preceeding 5 Years at the Borough Level"
       )
```

**Relationship Between Logged Adult Homeless Population and Logged Average YOY Percent Change in Completed Housing Related Construction Projects for Preceeding 5 Years at the Borough Level**
``` {r agg_log_reg1b, echo=FALSE, warning=FALSE, message=FALSE}

## Ground Syringes to Housing SUm
agg_lm1b <- lm(log(avg_homeless_count+1) ~ log(avg_yoy_change+1), data = data_final_boro_nonulls)
print(summary(agg_lm1b))
```

```{r agg_logreg1b_plot, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data = data_final_boro_nonulls, aes(x = log(avg_yoy_change+1), y = log(avg_homeless_count+1))) +
geom_point() +
stat_smooth(method = "lm", se = FALSE)+
    labs(x="Logged Avg. YOY Pct Change in Completed Housing Rltd Constr. Projects for Preceeding 5 Years",
       y="Logged Annual Average Adult Homeless Pppulation",
       # title="Relationship Between Logged Adult Homeless Population and Logged Average YOY Percent Change in Completed Housing Related Construction Projects for Preceeding 5 Years at the Borough Level"
       )
```

#### QUESTION 2: Does the size of the Adult Homeless Population influence the Number of Used Syringes recovered by NYC Parks? 

**Relationship Between Adult Homeless Population and Total Number of Used Syringes Collected at the Borough Level**
``` {r agg_reg2, echo=FALSE, warning=FALSE, message=FALSE}

## Ground Syringes to Avg. Annual Housing Increase
agg_m2 <- lm(total_syringes ~ avg_homeless_count, data = data_final_boro_nonulls)
print(summary(agg_m2))
```

```{r agg_reg2_plot, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data = data_final_boro_nonulls, aes(x = avg_homeless_count, y = total_syringes)) +
geom_point() +
stat_smooth(method = "lm", se = FALSE)+
      labs(x="Total Number of Used Syringes Collected",
       y="Annual Average Adult Homeless Pppulation",
       # title="Relationship Between Adult Homeless Population and Total Number of Used Syringes Collected at the Borough Level"
       )
```

**Relationship Between Logged Adult Homeless Population and Logged Total Number of Used Syringes Collected at the Borough Level**
``` {r agg_logreg2, , echo=FALSE, warning=FALSE, message=FALSE}

## Ground Syringes to Avg. Annual Housing Increase
agg_lm2 <- lm(log(total_syringes+1) ~ log(avg_homeless_count+1), data = data_final_boro_nonulls)
print(summary(agg_lm2))
```

```{r agg_logreg2_plot, , echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data = data_final_boro_nonulls, aes(x = log(avg_homeless_count+1), y = log(total_syringes+1))) +
geom_point() +
stat_smooth(method = "lm", se = FALSE)+
        labs(x="Logged Total Number of Used Syringes Collected",
       y="Logged Annual Average Adult Homeless Pppulation",
       # title="Relationship Between Logged Adult Homeless Population and Logged Total Number of Used Syringes Collected at the Borough Level"
       )
```


## Conclusion

# CHECK FOR SKEWS!!!!!


Of the 12 different attempts at linear regressions to identify strong relationships between varuables, only 5 had statistically significant results. 
- Housing projects focussing on luxury units in trendy areas? Not afforable housing. This may exaplin the direct relationship at the CD level.  This may be why when agg up to borough level we get the expected relationship with YOY change and the homeless population. We also get the expected relationship with the syringes and the homeless, so its possible that the granualr geog limited the relationships. 

We would want more years of data.

LOGGING

**All Regression Attempt Results**

| Regression Analysis                                      | Geography Level      | Dependent Variable          | Independent Variable          | R-squared | Adjusted R-squared | P-value          |
|---------------------------------------------------------|----------------------|-----------------------------|-------------------------------|-----------|--------------------|------------------|
| Adult Homeless Pop. & Total Housing Proj. Prev 5 Yrs.  | Community District   | avg_homeless_count          | housingsum                    | 0.2854    | 0.2807             | 1.15e-12         |
| Adult Homeless Pop. & Total Housing Proj. Prev 5 Yrs. (Logged) | Community District   | log(avg_homeless_count + 1) | log(housingsum + 1)           | 0.2725    | 0.2676             | 4.54e-12         |
| Adult Homeless Pop. & Avg. YOY Pct. Chg. Housing Proj. Prev 5 Yrs. | Community District   | avg_homeless_count          | prev5_avg_yoy_change          | 0.00018   | -0.0064            | 0.8694           |
| Adult Homeless Pop. & Avg. YOY Pct. Chg. Housing Proj. Prev 5 Yrs. (Logged) | Community District   | log(avg_homeless_count + 1) | log(prev5_avg_yoy_change + 1) | 0.00114   | -0.0077            | 0.7203           |
| Adult Homeless Pop. & Total Recovered Syringe Count     | Community District   | total_syringe_ests          | avg_homeless_count            | 0.00081   | -0.0058            | 0.7268           |
| Adult Homeless Pop. & Total Recovered Syringe Count (Logged) | Community District   | log(total_syringe_ests + 1) | log(avg_homeless_count + 1)   | 0.1096    | 0.1037             | 2.91e-05         |
| Adult Homeless Pop. & Total Housing Proj. Prev 5 Yrs.  | Borough              | avg_homeless_count          | housingsum                    | 0.3538    | 0.3290             | 0.00084          |
| Adult Homeless Pop. & Total Housing Proj. Prev 5 Yrs. (Logged) | Borough              | log(avg_homeless_count + 1) | log(housingsum + 1)           | 0.2725    | 0.2676             | 4.54e-12         |
| Adult Homeless Pop. & Avg. YOY Pct. Chg. Housing Proj. Prev 5 Yrs. | Borough              | avg_homeless_count          | avg_yoy_change                | 0.1948    | 0.1639             | 0.0187           |
| Adult Homeless Pop. & Avg. YOY Pct. Chg. Housing Proj. Prev 5 Yrs. (Logged) | Borough              | log(avg_homeless_count + 1) | log(avg_yoy_change + 1)       | 0.1922    | 0.1497             | 0.0468           |
| Adult Homeless Pop. & Total Recovered Syringe Count     | Borough              | total_syringes              | avg_homeless_count            | 0.1189    | 0.0850             | 0.0724           |
| Adult Homeless Pop. & Total Recovered Syringe Count (Logged) | Borough              | log(total_syringes + 1)     | log(avg_homeless_count + 1)   | 0.2189    | 0.1889             | 0.0120           |


**Statistically Significant Results**

| Regression Analysis                                      | Geography Level      | Dependent Variable          | Independent Variable          | R-squared | Adjusted R-squared | P-value          |
|---------------------------------------------------------|----------------------|-----------------------------|-------------------------------|-----------|--------------------|------------------|
| Adult Homeless Pop. & Total Housing Proj. Prev 5 Yrs.  | Community District   | avg_homeless_count          | housingsum                    | 0.2854    | 0.2807             | 1.15e-12         |
| Adult Homeless Pop. & Total Housing Proj. Prev 5 Yrs. (Logged) | Community District   | log(avg_homeless_count + 1) | log(housingsum + 1)           | 0.2725    | 0.2676             | 4.54e-12         |
| Adult Homeless Pop. & Avg. YOY Pct. Chg. Housing Proj. Prev 5 Yrs. | Borough              | avg_homeless_count          | avg_yoy_change                | 0.1948    | 0.1639             | 0.0187           |
| Adult Homeless Pop. & Avg. YOY Pct. Chg. Housing Proj. Prev 5 Yrs. (Logged) | Borough              | log(avg_homeless_count + 1) | log(avg_yoy_change + 1)       | 0.1922    | 0.1497             | 0.0468           |
| Adult Homeless Pop. & Total Recovered Syringe Count (Logged) | Borough              | log(total_syringes + 1)     | log(avg_homeless_count + 1)   | 0.2189    | 0.1889             | 0.0120           |

## References

^1^ "Home Prices, Wages, and Rent: Harvard Report Reveals Housing Trends." NPR, 20 June 2024, [https://www.npr.org/2024/06/20/nx-s1-5005972/home-prices-wages-paychecks-rent-housing-harvard-report#:~:text=In%20past%20decades%2C%20it%20was,Philly%20is%20giving%20renters%20cash](). Accessed 7 Dec. 2024.
