---
title: "Week5_Data607"
author: "John Ferrara"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading in the Library


```{r importing}
library(dplyr)
library(tidyr)
library(tidyverse)
```

## Looking at the data in the dplyr package


```{r pressure, echo=FALSE}
(data(package="dplyr")$results)
```
## Reading in the created CSV file 

``` {r reading_in}
path = path.expand("~/OneDrive/Desktop/CUNY_SPS_Materials/DATA607 - DATA ACQ/Week5/flight_data.csv")
print(path)

df = read.csv(path)
df

```

## Working to Transform the data in order to plot and compare the delays

``` {r transform}
## First parsing out the delayed and the ontime using filter function in dplyr
ontime <- df %>% filter(Status =="on time")
delayed <- df %>% filter(Status =="delayed")

## Using Tidyr  to pivot each one longways for analysis

##Citation: https://tidyr.tidyverse.org/reference/pivot_longer.html
long_ontime <- pivot_longer(ontime,
             cols = c("Los.Angeles","Phoenix","San.Diego","San.Francisco","Seattle"), names_to = "arrival_city", values_to ="instance_count")
long_delay <- pivot_longer(delayed,
             cols = c("Los.Angeles","Phoenix","San.Diego","San.Francisco","Seattle"), names_to = "arrival_city", values_to ="instance_count")

#Concatenatining them together for one coherent DF 
# CITATIOn: https://dplyr.tidyverse.org/reference/bind_rows.html
long_flights <- bind_rows(long_ontime,long_delay)
long_flights

```
## Transformation Part 2 (Alaska)
```{r transform1}
## Mean and Median for Alaska Airlines
alaska_del <- long_flights %>%
  filter(Airline=='ALASKA' & Status=='delayed') %>%
  summarize(
    total_del_count = sum(instance_count),
    avg_del_count = mean(instance_count),
    med_del_count = median(instance_count)
  ) 

alaska_del$airline = "ALASKA"
  
alaska_ot <-long_flights %>%
  filter(Airline=='ALASKA' & Status=='on time')%>%
  summarize(
    total_ot_count = sum(instance_count),
    avg_ot_count = mean(instance_count),
    med_ot_count = median(instance_count)
  ) 
alaska_ot$airline = "ALASKA"

alaska <- merge(alaska_del, alaska_ot, by = "airline")
alaska$pct_del <- (alaska$total_del_count / (alaska$total_del_count + alaska$total_ot_count))*100

alaska
```

##  Transformation Part 2 (AMWEST)
``` {r transform2}
# Mean and Median for AMWEST Airlines
## Mean and Median for Alaska Airlines
amwest_del <- long_flights %>%
  filter(Airline=='AMWEST' & Status=='delayed') %>%
  summarize(
    total_del_count = sum(instance_count),
    avg_del_count = mean(instance_count),
    med_del_count = median(instance_count)
  ) 

amwest_del$airline = "AMWEST"

amwest_ot <-long_flights %>%
  filter(Airline=='AMWEST' & Status=='on time')%>%
  summarize(
    total_ot_count = sum(instance_count),
    avg_ot_count = mean(instance_count),
    med_ot_count = median(instance_count)
  ) 
amwest_ot$airline = "AMWEST"

amwest <- merge(amwest_del, amwest_ot, by = "airline")
amwest$pct_del <- (amwest$total_del_count / (alaska$total_del_count + amwest$total_ot_count))*100

amwest

#Combining both descriptive wide dfs for analysis
 stat_table <- bind_rows(alaska,amwest)
 stat_table
```

## Analysis

**When all flights are considered Alaska airlines has a total of 13.27% of arrival delays, which is a bit more than the 11.34 percent arrival delays of AMWest. However, when looking at the count of delays by city on the second plot, the arrival delays vary widely depending on location. The best airline with the lesser amount of delays for arrivals in Los Angeles, Phoenix, and San Diego is Alaska airlines. While AMwest is the best airline for arrivals of the two for in Seatle. Arrival delays are San Francisco are more evenly distributed between the airlines, but AMWEST has a little more of them.**
```{r analysis}
#Plotting the main overview stats for relative percent delays performamce
ggplot(data = stat_table, aes(x=airline, y=pct_del))+
  geom_bar(stat = "identity")

#Looking at the long df again
head(long_flights)

ggplot(data = long_flights%>%filter(Status=='delayed'),aes(x=arrival_city,y=instance_count, fill=Airline))+
  geom_bar(stat = "identity")+
  geom_text(aes(label=instance_count),position = position_dodge(width = 0.8))

#Note the labels are off, but can be assumed based on the chart colors and sizes. 

```

## Conclusion
**Overall, the airlines are very similar in performance when looked at as a percetage of arrival delays compared to all of their respective flights. Alaska airlines does have a few more arriaval delays than AMWEST. However, when looked at with more of a geographic lense, one may want to select Alaska airlines if going to Los Angeles, Phoenix or San Diego because AMWEST has many more arrival delays in those airports. If traveling to Seattle, one may want to fly with AMWEST. FOr San Francisco, both airlines seem to perform similarly.  **